### 2018-05-02 proposal/proposal.tex
Update proposal with final feedback

Another property that all of the aforementioned global land cover maps share is that each pixel is said to belong to a certain land cover class; the classification is based on classes with \textit{discrete} labels. In reality, land cover does not follow a grid pattern, and a large proportion of pixels in a moderate resolution remote sensing image cover an area that includes more than a single land cover class. This is a problem especially in heterogeneous areas with small patch sizes. Classification with discrete labels can only express the dominant land cover type in a pixel, any information about other land cover types in the same pixel is lost. In contrast, producing a map of \textit{fractions} of land cover classes allows for expressing the proportion of each land cover class in each pixel, which in turn eliminates the need for mosaic classes and makes the legend easier to interpret. Fractional land cover maps can also be reduced to discretely labelled maps, allowing users to define their own rules and thresholds for doing so, thus taking into accout which classes the user is interested in \citep{tsendbazar_integrating_2017}. In addition, coarser resolution fractional land cover mapping reduces the uncertainty of classification, since the requirement of knowing the specific location of a particular land cover class is relaxed. This addresses the needs of the scientific community, where models do not require high spatial resolution, but require higher accuracy \citep{bontemps_revisiting_2012}.

Fractional classification (also known as fuzzy classificaiton) has been popular in the past, when only coarse resolution imagery was available \citep{foody_approaches_1996, adams_classification_1995, amo_spectral_2002}. In recent years, fractional classification has been gaining support once again, as a method to better represent land cover \citep{gessner_estimating_2013}, but the suitability for global land cover mapping as well as the issue of updating fractional land cover maps has been largely unexplored up to now. Current global land cover products only include cover fractions for specific land cover classes, such as forests \citep{hansen_global_2003}, rather than covering the full range of land cover classes, and they do not yet distinguish between different tree cover types.

@@

\begin{enumerate}
 \item There is a variety of machine learning algorithms that are able to produce information about the fractions of land cover classes, but some are not suitable for continental or global scale classification due to their computational intensity, low accuracy (overfitting), and/or ability to deal with non-normal or collinear variables.
 \item Extracting temporal features from time series of satellite data without consideration of land cover change occurring during the time period leads to the temporal metrics being a mix between several land cover classes. Break detection in time series can be used to detect periods of stable land cover.
 \item Differences between land cover maps produced anually are often spurious, rather than reflecting actual changes on the ground. This issue can be eliminated by making the methods of producing the maps more conservative, rather than presuming that either of the two maps is error-free.
 \item Discrete land cover classes cannot express gradual changes from one class to another, only abrupt changes affecting an entire pixel. Using class fractions would allow for expressing such changes.
\end{enumerate}

@@

\begin{enumerate}[label=RO\arabic*]
 \item \label{RO1} To compare the suitability of machine learning algorithms for performing continental-scale fractional land cover mapping.
 \item \label{RO2} To improve the robustness of time series-derived temporal metrics used in land cover mapping by performing land cover change detection on dense time series.
 \item \label{RO3} To compare methods for creating anually consistent and stable global land cover change maps with discrete class labels. % And error propagation/uncertainty?
 \item \label{RO4} To investigate the methods of representing land cover changes between years using a set of fractional land cover maps.
\end{enumerate}

@@

These methods will be applied on time series data derived from the PROBA-V satellite archive (100 m resolution imagery at the equator) over the continent of Africa, and one case study site in Europe in order to test the applicability of the methods to a boreal forest gradient. Auxiliary data will also be included to serve as extra information for deriving the land cover fractions. The algorithms will be trained and validated on land cover sampling data that has been collected as part of the Copernicus Global Land Operations ``Vegetation and Energy'' (CGLOPS-1) project (see figure \ref{fig-sampling-fractions}, over 68000 sampling points in separate training and validation sets). Processing will be carried out on the Proba-V Mission Exploitation Platform (MEP), which allows direct access to Proba-V satellite imagery.

@@

@@

\centerline{
  \begin{minipage}[t]{0.9\linewidth}
    %\vspace{-2ex}
    \includegraphics[width=\linewidth]{figures/sampling-fractions}
    \captionof{figure}{Screenshot from an interpretation of a sampling pixel \citep{tsendbazar_validation_2018}. An area covering a Proba-V pixel (nominally 100×100 m) is subdivided into smaller polygons (nominally 10×10 m). These polygons are interpreted by an expert to obtain fractional land cover validation data for each pixel. The acquisition date of the imagery used for interpretation is recorded as well. Over 68000 such sampling pixels have been interpreted so far as part of the Copernicus Global Land Operations ``Vegetation and Energy'' (CGLOPS-1) project.}
    \label{fig-sampling-fractions}
    \bigskip
  \end{minipage}
}

@@

\begin{mdframed}[style=table,frametitle=\textbf{8. TIME TABLE OF THE PROJECT AND WORK PROGRAMME}]
%\resizebox{0.7\textwidth}{!}{
\centerline{
    \begin{ganttchart}[hgrid, vgrid, y unit chart=0.9cm]{1}{16}
      \gantttitle{{\tiny 2017}}{1} \gantttitlelist{2018,...,2020}{4} \gantttitle{2021}{3} \\
      \gantttitle[title label node/.append style={below left=-7pt and -6pt}]{Quarters:\quad4}{1} \gantttitlelist{1,...,4}{1} \gantttitlelist{1,...,4}{1} \gantttitlelist{1,...,4}{1} \gantttitlelist{1, ..., 3}{1} \\
      \ganttbar{TSP and proposal writing}{1}{3} \\
      \ganttgroup{Research objective 1}{3}{5} \\
      \ganttbar{Fractional mapping over Africa}{3}{5} \\
      \ganttbar{Paper writing (RO1)}{4}{5} \\
      \ganttgroup{Research objective 2}{2}{7} \\
      \ganttbar{Breakpoint detection}{2}{3} \ganttbar{}{6}{7} \\
      \ganttbar{Paper writing (RO2)}{6}{7} \\
      \ganttgroup{Research objective 3}{4}{10} \\
      \ganttbar{Land cover change modelling, discrete}{4}{4} \ganttbar{}{8}{10} \\
      \ganttbar{Paper writing (RO3)}{9}{10} \\
      \ganttgroup{Research objective 4}{11}{13} \\
      \ganttbar{Land cover change modelling, fractional}{11}{13} \\
      \ganttbar{Paper writing (RO4)}{12}{13} \\
      \ganttgroup{Thesis finalisation}{14}{16} \\
      \ganttbar{Synthesis}{14}{15} \\
      \ganttmilestone{Thesis defence}{16}
    \end{ganttchart}
}
%}
\end{mdframed}

>>>>

Another property that all of the aforementioned global land cover maps share is that each pixel is said to belong to a certain land cover class; the classification is based on classes with \textit{discrete} labels. In reality, land cover does not follow a grid pattern, and a large proportion of pixels in a moderate resolution remote sensing image cover an area that includes more than a single land cover class. This is a problem especially in heterogeneous areas with small patch sizes. Classification with discrete labels can only express the dominant land cover type in a pixel, any information about other land cover types in the same pixel is lost. In contrast, producing a map of \textit{fractions} of land cover classes allows for expressing the proportion of each land cover class in each pixel, which in turn eliminates the need for mosaic classes and makes the legend easier to interpret. Fractional land cover maps can also be reduced to discretely labelled maps, allowing users to define their own rules and thresholds for doing so, thus taking into account which classes the user is interested in \citep{tsendbazar_integrating_2017}. In addition, coarser resolution fractional land cover mapping reduces the uncertainty of classification, since the requirement of knowing the specific location of a particular land cover class is relaxed. This addresses the needs of the scientific community, where models do not require high spatial resolution, but require higher accuracy \citep{bontemps_revisiting_2012}.

Fractional classification (also known as fuzzy classification) has been popular in the past, when only coarse resolution imagery was available \citep{foody_approaches_1996, adams_classification_1995, amo_spectral_2002}. In recent years, fractional classification has been gaining support once again, as a method to better represent land cover \citep{gessner_estimating_2013}, but the suitability for global land cover mapping as well as the issue of updating fractional land cover maps has been largely unexplored up to now. Current global land cover products only include cover fractions for specific land cover classes, such as forests \citep{hansen_global_2003}, rather than covering the full range of land cover classes, and they do not yet distinguish between different tree cover types.

@@

\begin{enumerate}
 \item There is a variety of machine learning algorithms that are able to produce information about the fractions of land cover classes, but some are not suitable for continental or global scale classification due to their computational intensity, low accuracy (overfitting), and/or ability to deal with non-normal or collinear variables.
 \item Extracting temporal features from time series of satellite data without consideration of land cover change occurring during the time period leads to the temporal metrics being a mix between several land cover classes. Break detection in time series can be used to detect periods of stable land cover.
 \item Differences between land cover maps produced annually are often spurious, rather than reflecting actual changes on the ground. This issue can be eliminated by making the methods of producing the maps more conservative, rather than presuming that either of the two maps is error-free.
 \item Discrete land cover classes cannot express gradual changes from one class to another, only abrupt changes affecting an entire pixel. Using class fractions would allow for expressing such changes.
\end{enumerate}

@@

\begin{enumerate}[label=RO\arabic*]
 \item \label{RO1} To compare the suitability of machine learning algorithms for performing continental-scale fractional land cover mapping.
 \item \label{RO2} To improve the robustness of time series-derived temporal metrics used in land cover mapping by performing land cover change detection on dense time series.
 \item \label{RO3} To compare methods for creating annually consistent and stable global land cover change maps with discrete class labels. % And error propagation/uncertainty?
 \item \label{RO4} To investigate the methods of representing land cover changes between years using a set of fractional land cover maps.
\end{enumerate}

@@

These methods will be applied on time series data derived from the PROBA-V satellite archive (100 m resolution imagery at the equator) over the continent of Africa, and one case study site in Europe in order to test the applicability of the methods to a boreal forest gradient. Auxiliary data will also be included to serve as extra information for deriving the land cover fractions. The algorithms will be trained and validated on land cover sampling data that has been collected as part of the Copernicus Global Land Operations ``Vegetation and Energy'' (CGLOPS-1) project (see figure \ref{fig-sampling-fractions}, over 25000 sampling points over Africa in separate training and validation sets). Processing will be carried out on the Proba-V Mission Exploitation Platform (MEP), which allows direct access to Proba-V satellite imagery.

@@

These methods will also be compared to Markov Random Fields, which is a post-classification technique for reducing changes in labels. It takes both the spatial and temporal variability around each pixel into account and can use rules to define how easily a class may transition to another given class. This approach is used in some existing land cover maps (e.g. \citet{cai_enhancing_2014}).

@@

\centerline{
  \begin{minipage}[t]{0.9\linewidth}
    %\vspace{-2ex}
    \includegraphics[width=\linewidth]{figures/sampling-fractions}
    \captionof{figure}{Screenshot from an interpretation of a sampling pixel \citep{tsendbazar_validation_2018}. An area covering a Proba-V pixel (nominally 100×100 m) is subdivided into smaller polygons (nominally 10×10 m). These polygons are interpreted by an expert to obtain fractional land cover validation data for each pixel. The acquisition date of the imagery used for interpretation is recorded as well. Over 68000 such sampling pixels have been interpreted globally so far as part of the Copernicus Global Land Operations ``Vegetation and Energy'' (CGLOPS-1) project.}
    \label{fig-sampling-fractions}
    \bigskip
  \end{minipage}
}

@@

\begin{mdframed}[style=table,frametitle=\textbf{8. TIME TABLE OF THE PROJECT AND WORK PROGRAMME}]
%\resizebox{0.7\textwidth}{!}{
\centerline{
    \begin{ganttchart}[hgrid, vgrid, y unit chart=0.9cm]{1}{16}
      \gantttitle{{\tiny 2017}}{1} \gantttitlelist{2018,...,2020}{4} \gantttitle{2021}{3} \\
      \gantttitle[title label node/.append style={below left=-7pt and -6pt}]{Quarters:\quad4}{1} \gantttitlelist{1,...,4}{1} \gantttitlelist{1,...,4}{1} \gantttitlelist{1,...,4}{1} \gantttitlelist{1, ..., 3}{1} \\
      \ganttbar{TSP and proposal writing}{1}{3} \\
      \ganttgroup{Research objective 1}{3}{5} \\
      \ganttbar{Fractional mapping over Africa}{3}{5} \\
      \ganttbar{Paper writing (RO1)}{4}{5} \\
      \ganttgroup{Research objective 2}{2}{7} \\
      \ganttbar{Breakpoint detection}{2}{3} \ganttbar{}{6}{7} \\
      \ganttbar{Paper writing (RO2)}{6}{7} \\
      \ganttgroup{Research objective 3}{4}{10} \\
      \ganttbar{Land cover change modelling, discrete}{4}{4} \ganttbar{}{8}{10} \\
      \ganttbar{Paper writing (RO3)}{9}{10} \\
      \ganttgroup{Research objective 4}{11}{14} \\
      \ganttbar{Land cover change modelling, fractional}{11}{14} \\
      \ganttbar{Paper writing (RO4)}{13}{14} \\
      \ganttgroup{Thesis finalisation}{14}{16} \\
      \ganttbar{Synthesis}{14}{15} \\
      \ganttmilestone{Thesis defence}{16}
    \end{ganttchart}
}
%}
\end{mdframed}
